---
title: "Introduction to R"
output:
html_notebook:
  toc: true
  toc_float: true
---

```{r include = FALSE}
# install.packages("tidyverse")
# install.packages("plotly")
library(tidyverse)
library(plotly)

```


# Introduction

First of all, let's examine the **cars** dataset. Display the first 6 rows.

```{r echo=FALSE}

head(cars)

```

This is the entire dataset.

```{r echo=FALSE}

cars

```

# LaTeX code

$\frac{a+b}{c+d}$

$$\lim\limits_{x \to \infty} \exp(-x) = 0$$

# Exploratory Data Analysis

Let's investigate the size and dimensionality of our dataset.

```{r echo=FALSE}

n_rows <- nrow(cars)

n_cols <- ncol(cars)

n_rows

n_cols

```

The size is `r n_rows` and the dimensionality is `r n_cols`.

## Variation within variables

Let's investigate our dataset using a 5 number summary.

```{r echo=FALSE}

summary(cars)

```

Let's visualise our distributions using boxplots. This is using base R graphics.

```{r echo=FALSE}
boxplot(cars$speed)

boxplot(cars$dist)
```

This is a ggplot alternative. 

```{r}

cars %>%
  ggplot(aes(y=speed)) +
  geom_boxplot()
  # geom_histogram(binwidth=5)

ggplotly()

cars %>% 
  ggplot(aes(y=dist)) +
  geom_boxplot()

ggplotly()
```

This is base R code for scatterplot.

```{r}
plot(cars)
```


This is a ggplot scatterplot

```{r}
xy_plot <- cars %>%
  ggplot(aes(x=speed, y=dist)) +
  geom_point() +
  theme_bw()

ggplotly(xy_plot)

```


```{r}
round(cor(cars$speed,cars$dist),2)
```






